\mainmatter

# Introduction {#intro}

<!-- - Gap between BioInfo and BioC -->
<!--   + Heavy BioInfo tools and Pipelines beyond BioC -->

<!-- - Tools to build pipelines and generate reproducible reports are required  -->
<!--   + Standardization of BioInfo Pipelines for best practice -->
<!--   + Portable pipelines -->
<!--   + Reproducible analysis and report -->
<!--   + Easy to use in different platforms -->

## Common Workflow Language
"The Common Workflow Language (CWL) is a specification for describing
analysis workflows and tools in a way that makes them portable and
scalable across a variety of software and hardware environments, from
workstations to **cluster, cloud, and high performance computing (HPC)**
environments."

<https://www.commonwl.org/>

- To wrap tool and workflow parameters in a standard format

- Capable of invoking tools from Docker containers

- Widely used...

## First example
The main class and constructor function is `cwlParam`, which wraps a command line tool and its parameters in a `cwlParam` object. Let's start with a simple example, `echo hello world`.

First, we load the package and define the input parameter for "echo", a string without a prefix. Just an `id` option required here:
```{r}
input1 <- InputParam(id = "sth")
```

Second, we create a `cwlParam` object with `baseCommand` for the command to execute and `InputParamList` for the input parameters.
```{r}
echo <- cwlParam(baseCommand = "echo", inputs = InputParamList(input1))
```

Now we have a command object to run. Let's send a string "Hello World!" to the object. Without defining the outputs, it will stream standard output to a temporary file by default.
```{r}
echo$sth <- "Hello World!"
echo
```

## Test run
The function `runCWL` is used to run the CWL object by invoking the python tool `cwltool`. The return will be a list including the command executed, temporary output and logs. The output directory is the current folder by default, but it can be changed by setting `outdir` option. All standard out and standard error streams can also be printed by setting `stderr = ""`.
```{r}
r1 <- runCWL(echo, outdir = tempdir())
r1
```

Here we can check the output to ensure the code did what we expected.
```{r}
r1$output
readLines(r1$output)
```

The executed command was returned in the result list. It shows the command that we have defined to execute.
```{r}
r1$command
```

The log shows the details of how the `cwltool` works with CWL scripts.
```{r}
r1$log
```
The `runCWL` generated two scripts with the default `tempfile` prefix, the tool wrapper CWL file and the input YML file. The `cwltool` parses the two scripts and translates them into the command shown before. The output is not defined in the `cwlParam` object, so the command output was returned to __stdout__ by default.
