# Components
## Input Parameters
### Essential Input parameters
For the input parameters, we usually need to define three options, *id*, *type*, and *prefix*. The type can be *string*, *int*, *long*, *float*, *double*, and so on. More detail can be found at: <https://www.commonwl.org/v1.0/CommandLineTool.html#CWLType>.

A `InputParam` constructor is used to define a list of input objects for the command tool, such as "prefix" for the parameter flags and "position" for the paramter orders. More descriptions are available in the CWL specification <https://www.commonwl.org/v1.0/CommandLineTool.html#CommandLineBinding>.

Here is an example from the CWL user guide(<http://www.commonwl.org/user_guide/03-input/>). We defined the `echo` with different type of input parameters by `InputParam`, and the `stdout` option can be used to caputre the standard output stream into a file:

```{r}
e1 <- InputParam(id = "flag", type = "boolean", prefix = "-f")
e2 <- InputParam(id = "string", type = "string", prefix = "-s")
e3 <- InputParam(id = "int", type = "int", prefix = "-i")
e4 <- InputParam(id = "file", type = "File", prefix = "--file=", separate = FALSE)
echoA <- cwlParam(baseCommand = "echo",
                  inputs = InputParamList(e1, e2, e3, e4),
                  stdout = "output.txt")
```

Then we give it a try by setting values for the inputs:
```{r}
echoA$flag <- TRUE
echoA$string <- "Hello"
echoA$int <- 1

tmpfile <- tempfile()
write("World", tmpfile)
echoA$file <- tmpfile

r2 <- runCWL(echoA, outdir = tempdir())
r2$command
```

### Array Inputs
Taking a similar example to the CWL user guide described above, we can define three different type of array as inputs: 
```{r}
a1 <- InputParam(id = "A", type = "string[]", prefix = "-A")
a2 <- InputParam(id = "B",
                 type = InputArrayParam(items = "string",
                                        prefix="-B=", separate = FALSE))
a3 <- InputParam(id = "C", type = "string[]", prefix = "-C=",
                 itemSeparator = ",", separate = FALSE)
echoB <- cwlParam(baseCommand = "echo",
                 inputs = InputParamList(a1, a2, a3))
```

We then set values for the three inputs:
```{r}
echoB$A <- letters[1:3]
echoB$B <- letters[4:6]
echoB$C <- letters[7:9]
echoB
```

Now we can check whether the command behaves as we expected:
```{r}
r3 <- runCWL(echoB, outdir = tempdir())
r3$command
```

## Output Parameters
### Capturing Output
The outputs, similar to the inputs, is a list of output parameters. Three options, *id*, *type* and *glob*, can be defined. The *glob* option is used to define a pattern to find files relative to the output directory.

Here is an example to unzip a compressed `gz` file. First, we generate a compressed R script file:
```{r}
zzfil <- file.path(tempdir(), "sample.R.gz")
zz <- gzfile(zzfil, "w")
cat("sample(1:10, 5)", file = zz, sep = "\n")
close(zz)
```

We then define a `cwlParam` object to use "gzip" to uncompress an input file:
```{r}
ofile <- "sample.R"
z1 <- InputParam(id = "uncomp", type = "boolean", prefix = "-d")
z2 <- InputParam(id = "out", type = "boolean", prefix = "-c")
z3 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
gz <- cwlParam(baseCommand = "gzip",
               inputs = InputParamList(z1, z2, z3),
               outputs = OutputParamList(o1),
               stdout = ofile)
```

Now the `gz` object can be used to uncompress the previously generated compressed file:
```{r}
gz$uncomp <- TRUE
gz$out <- TRUE
gz$zfile <- zzfil
r4 <- runCWL(gz, outdir = tempdir())
r4$output
```

Or we can use `arguments` to set some default parameters:
```{r}
z1 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
Gz <- cwlParam(baseCommand = "gzip",
               arguments = list("-d", "-c"),
               inputs = InputParamList(z1),
               outputs = OutputParamList(o1),
               stdout = ofile)
Gz
Gz$zfile <- zzfil
r4a <- runCWL(Gz, outdir = tempdir())
```

To make it for general usage, we can define a pattern with javascript to *glob* the output, which requires `node` to be installed in your system PATH:
```{r}
pfile <- "$(inputs.zfile.path.split('/').slice(-1)[0].split('.').slice(0,-1).join('.'))"
```

Or we can directly use the CWL built in file property, `nameroot`:
```{r}
pfile <- "$(inputs.zfile.nameroot)"
o2 <- OutputParam(id = "rfile", type = "File", glob = pfile)
req1 <- list(class = "InlineJavascriptRequirement")
GZ <- cwlParam(baseCommand = c("gzip", "-d", "-c"),
               requirements = list(), ## assign list(req1) if node installed.
               inputs = InputParamList(z1),
               outputs = OutputParamList(o2),
               stdout = pfile)
GZ$zfile <- zzfil
r4b <- runCWL(GZ, outdir = tempdir())
```

### Array Outputs

We can also capture multiple output files with the `glob` pattern:
```{r}
a <- InputParam(id = "a", type = InputArrayParam(items = "string"))
b <- OutputParam(id = "b", type = OutputArrayParam(items = "File"), glob = "*.txt")
touch <- cwlParam(baseCommand = "touch", inputs = InputParamList(a), outputs = OutputParamList(b))
touch$a <- c("a.txt", "b.gz", "c.txt")
r5 <- runCWL(touch, outdir = tempdir())
r5$output
```

### Standard output

Usually, the `stdout` option is a string or an expression of output
file name from the command tool. The command's standard output stream
will be captured into a file written to the designated output
directory. When the `stdout` field is defined, an output parameter
with the type of "stdout" should be also assigned with no
"outputBinding" set.

An example for command tool "cat" is defined with `stdout` field in
the output, with the name passed from the input parameter "p2": 

```{r}
## define Cat
p1 <- InputParam(id = "infiles", type = "File[]")
p2 <- InputParam(id = "outfile", type = "string",
                 default = "catout.txt", position = -1)
Cat <- cwlParam(baseCommand = "cat",
                inputs = InputParamList(p1, p2),
                stdout = "$(inputs.outfile)")
## assign inputs
afile <- file.path(tempdir(), "a.txt")
bfile <- file.path(tempdir(), "b.txt")
write("a", afile)
write("b", bfile)
Cat$infiles <- list(afile, bfile)
## run
r6 <- runCWL(Cat, outdir = tempdir())
r6$command
```

```{block2, type='tips'}
In this example, we used the parameter "p2" to pass the name to the
standard output.

In the `InputParam` of "p2", the position is assigned to a negative
value (-1), which means the parameters will not be used in the command
and only uses for passing variable. To write the "Cat" tool to a CWL
file, the "inputBinding" field will be skipped for this parameter.

```
