\mainmatter

# Get started 

`cwlProcess` is the main constructor function to wrap a command line
tool into an _R_ tool as a `cwlProcess` object (S4 class).  Let's
start with a simple example to wrap the `echo` command and execute
`echo hello world` in _R_.

First, we need to define the input parameter for the base command
`echo`, here it is a string without a prefix. An `id` argument is
required here.

```{r}
input1 <- InputParam(id = "sth")
```

Second, we can construct a `cwlProcess` object by specifying the
`baseCommand` for the command line tool, and `InputParamList` for the
input parameters. 
```{r}
echo <- cwlProcess(baseCommand = "echo", inputs = InputParamList(input1))
```

Now we have converted the command line tool `echo` into an _R_ tool:
an _R_ object of class `cwlProcess` with the name of `echo`. We can
take a look at the this _R_ object and use some utility functions to
extract specific information.

```{r}
echo
class(echo)
cwlClass(echo)
cwlVersion(echo)
baseCommand(echo)
inputs(echo)
outputs(echo)
```

The `inputs(echo)` will show the value once it is assigned in next
step. Since we didn't define the outputs for this tool, it will stream
standard output to a temporary file by default.

The third step is to assign values (here is "Hello World!") for the
input parameters.

```{r}
echo$sth <- "Hello World!"
inputs(echo)
```

Now this _R_ version of command line tool `echo` is ready to be
executed. 

The function `runCWL` runs the tools in _R_ and returns a list of: 1)
actual command line that was executed, 2) filepath to the output, and
3) running logs. The output directory by default takes the working
directory, but can be specified in `outdir` argument.

```{r}
r1 <- runCWL(echo, outdir = tempdir())
r1
r1$command
readLines(r1$output)
r1$logs
```

Users can also have the log printed out by specifying `showLog = TRUE`.

```{r}
r1 <- runCWL(echo, outdir = tempdir(), showLog = TRUE)
```

A utility function `writeCWL` converts the `cwlProcess` object into 2
files: a `.cwl` file for the command and `.yml` file for the inputs,
which are the internal cwl files to be executed when `runCWL` is
invoked. The internal execution requires a `cwl-runner` (e.g.,
`cwltool`), which will be installed automatically with `runCWL`.

```{r}
writeCWL(echo)
```
