# Wrap command line tools

## Input Parameters

### Essential Input parameters

For the input parameters, three options need to be defined usually,
*id*, *type*, and *prefix*. The type can be *string*, *int*, *long*,
*float*, *double*, and so on. More detail can be found at:
<https://www.commonwl.org/v1.0/CommandLineTool.html#CWLType>.

Here is an example from [CWL user
guide](http://www.commonwl.org/user_guide/03-input/). Here we defined
an `echo` with different type of input parameters by `InputParam`. The
`stdout` option can be used to capture the standard output stream to a
file.

```{r}
e1 <- InputParam(id = "flag", type = "boolean", prefix = "-f")
e2 <- InputParam(id = "string", type = "string", prefix = "-s")
e3 <- InputParam(id = "int", type = "int", prefix = "-i")
e4 <- InputParam(id = "file", type = "File", prefix = "--file=", separate = FALSE)
echoA <- cwlProcess(baseCommand = "echo",
                  inputs = InputParamList(e1, e2, e3, e4),
                  stdout = "output.txt")
```

Then we can assign values for the input parameters.

```{r}
echoA$flag <- TRUE
echoA$string <- "Hello"
echoA$int <- 1

tmpfile <- tempfile()
write("World", tmpfile)
echoA$file <- tmpfile
```

```{r}
r2 <- runCWL(echoA, outdir = tempdir())
r2$command
```

The command shows the parameters work as we defined. The parameters
are in alphabetical orders by default, but can be modified by the
`position` argument in `InputParam` function.

### Array Inputs

A similar example to CWL user guide. We can define three different
type of array as inputs.

```{r}
a1 <- InputParam(id = "A", type = "string[]", prefix = "-A")
a2 <- InputParam(id = "B",
                 type = InputArrayParam(items = "string",
                                        prefix="-B=", separate = FALSE))
a3 <- InputParam(id = "C", type = "string[]", prefix = "-C=",
                 itemSeparator = ",", separate = FALSE)
echoB <- cwlProcess(baseCommand = "echo",
                 inputs = InputParamList(a1, a2, a3))
```

Then we can assign values for the three input parameters:

```{r}
echoB$A <- letters[1:3]
echoB$B <- letters[4:6]
echoB$C <- letters[7:9]
echoB
```

Now we can check whether the command behaves as we expected.

```{r}
r3 <- runCWL(echoB, outdir = tempdir())
r3$command
```

## Output Parameters

### Capturing Output

Similar to the input parameters, the output is a list of output
parameters. Three options *id*, *type* and *glob* can be defined. The
glob option is used to define a pattern to find files relative to the
output directory.

Here is an example to unzip a compressed `gz` file. First, we generate
a compressed _R_ script file.

```{r}
zzfil <- file.path(tempdir(), "sample.R.gz")
zz <- gzfile(zzfil, "w")
cat("sample(1:10, 5)", file = zz, sep = "\n")
close(zz)
```

Then we build a tool called `gz` (a `cwlProcess` object) to uncompress
a input file using 'gzip'.

```{r}
ofile <- "sample.R"
z1 <- InputParam(id = "uncomp", type = "boolean", prefix = "-d")
z2 <- InputParam(id = "out", type = "boolean", prefix = "-c")
z3 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
gz <- cwlProcess(baseCommand = "gzip",
               inputs = InputParamList(z1, z2, z3),
               outputs = OutputParamList(o1),
               stdout = ofile)
```

Now the `gz` is ready to uncompress the previous generated compressed
file.

```{r}
gz$uncomp <- TRUE
gz$out <- TRUE
gz$zfile <- zzfil
r4 <- runCWL(gz, outdir = tempdir())
r4$output
```

We can use `arguments` argument to modify some default parameters.

```{r}
z1 <- InputParam(id = "zfile", type = "File")
o1 <- OutputParam(id = "rfile", type = "File", glob = ofile)
Gz <- cwlProcess(baseCommand = "gzip",
               arguments = list("-d", "-c"),
               inputs = InputParamList(z1),
               outputs = OutputParamList(o1),
               stdout = ofile)
Gz
Gz$zfile <- zzfil
r4a <- runCWL(Gz, outdir = tempdir())
```

To make it for general usage, we can define a pattern with javascript
to glob the output, which requires `node` from 'nodejs' to be installed
in your system PATH.

```{r}
pfile <- "$(inputs.zfile.path.split('/').slice(-1)[0].split('.').slice(0,-1).join('.'))"
```

Or we can use the CWL built-in file property, `nameroot`, directly.

```{r}
pfile <- "$(inputs.zfile.nameroot)"
o2 <- OutputParam(id = "rfile", type = "File", glob = pfile)
req1 <- requireJS()
GZ <- cwlProcess(baseCommand = "gzip",
               arguments = list("-d", "-c"),
               requirements = list(), ## assign list(req1) if node installed.
               inputs = InputParamList(z1),
               outputs = OutputParamList(o2),
               stdout = pfile)
GZ$zfile <- zzfil
r4b <- runCWL(GZ, outdir = tempdir())
```

### Array Outputs

We can also capture multiple output files with `glob` pattern.

```{r}
a <- InputParam(id = "a", type = InputArrayParam(items = "string"))
b <- OutputParam(id = "b", type = OutputArrayParam(items = "File"),
                 glob = "*.txt")
touch <- cwlProcess(baseCommand = "touch", inputs = InputParamList(a),
                    outputs = OutputParamList(b))
touch$a <- c("a.txt", "b.log", "c.txt")
r5 <- runCWL(touch, outdir = tempdir())
r5$output
```

The 'touch' command generates three files, but the output only collects
two files with '.txt' suffix as defined in the `OutputParam` using the
'glob' option.

### Standard output

Usually, the `stdout` option is a string or an expression of output
file name from the command line tool. The command's standard output
stream will be captured into a file written to the designated output
directory. When the `stdout` field is defined, an output parameter
with the type of "stdout" should be also assigned with no
"outputBinding" set.

An example for command tool "cat" is defined with `stdout` field in
the output, with the name passed from the input parameter "p2": 

```{r}
## define Cat
p1 <- InputParam(id = "infiles", type = "File[]")
p2 <- InputParam(id = "outfile", type = "string",
                 default = "catout.txt", position = -1)
Cat <- cwlProcess(baseCommand = "cat",
                  inputs = InputParamList(p1, p2),
                  stdout = "$(inputs.outfile)")
```

```{r}
## assign values to inputs
afile <- file.path(tempdir(), "a.txt")
bfile <- file.path(tempdir(), "b.txt")
write("a", afile)
write("b", bfile)
Cat$infiles <- list(afile, bfile)
```

```{r}
## run the tool
r6 <- runCWL(Cat, outdir = tempdir())
r6$command
```

```{block2, type='tips'}
In this example, we used the parameter "p2" to pass the name to the
standard output.

In the `InputParam` of "p2", the position is assigned to a negative
value (-1), which means the parameters will not be used in the command
and only uses for passing variable. To write the "Cat" tool to a CWL
file, the "inputBinding" field will be skipped for this parameter.
```
