# Writing Pipeline

We can connect multiple tools together into a pipeline. Here is an
example to uncompress an R script and execute it with `Rscript`.

Here we define a simple `Rscript` tool without using docker.
```{r}
d1 <- InputParam(id = "rfile", type = "File")
Rs <- cwlProcess(baseCommand = "Rscript",
               inputs = InputParamList(d1))
Rs
```

Test run:
```{r}
Rs$rfile <- r4$output
tres <- runCWL(Rs, outdir = tempdir())
readLines(tres$output)
```

The pipeline includes two steps, decompressing with predefined
`cwlProcess` of `GZ` and compiling with `cwlProcess` of `Rs`. The
input file is a compressed file for the first "Uncomp" step.

```{r}
i1 <- InputParam(id = "cwl_zfile", type = "File")
s1 <- cwlStep(id = "Uncomp", run = GZ,
              In = list(zfile = "cwl_zfile"))
s2 <- cwlStep(id = "Compile", run = Rs,
              In = list(rfile = "Uncomp/rfile"))
```

In step 1 ('s1'), the pipeline runs the `cwlProcess` of `GZ`, where
the input `zfile` is defined in 'i1' with id of "cwl_zfile". In step 2
('s2'), the pipeline runs the `cwlProcess` of `Rs`, where the input
`rfile` is from the output of the step 1 ("Uncomp/rfile") using the
format of `<step>/<output>`.

The pipeline output will be defined as the output of the step 2
("Compile/output") using the format of `<step>/<output>` as shown
below.

```{r}
o1 <- OutputParam(id = "cwl_cout", type = "File",
                  outputSource = "Compile/output")
```

The `cwlWorkflow` function is used to initiate the pipeline by
specifying the `inputs` and `outputs`. Then we can simply use `+` to
connect all steps to build the final pipeline.

```{r}
cwl <- cwlWorkflow(inputs = InputParamList(i1),
                    outputs = OutputParamList(o1))
cwl <- cwl + s1 + s2
cwl
```

Let's run the pipeline.
```{r}
cwl$cwl_zfile <- zzfil
r7 <- runCWL(cwl, outdir = tempdir())
readLines(r7$output)
```

Tips: Sometimes, we need to adjust some arguments of certain tools in
a pipeline besides of parameter inputs. The function `arguments` can
help to modify arguments for a tool, tool in a pipeline, or even tool
in a sub-workflow. For example,

```{r}
arguments(cwl, step = "Uncomp") <- list("-d", "-c", "-f")
runs(cwl)$Uncomp
```


## Scattering pipeline
The scattering feature can specifies the associated workflow step or
subworkflow to execute separately over a list of input elements. To
use this feature, `ScatterFeatureRequirement` must be specified in the
workflow requirements. Different `scatter` methods can be used in the
associated step to decompose the input into a discrete set of
jobs. More details can be found at:
https://www.commonwl.org/v1.0/Workflow.html#WorkflowStep.

Here is an example to execute multiple R scripts. First, we need to
set the input and output types to be array of "File", and add the
requirements. In the "Compile" step, the scattering input is required
to be set with the `scatter` option.

```{r}
i2 <- InputParam(id = "cwl_rfiles", type = "File[]")
o2 <- OutputParam(id = "cwl_couts", type = "File[]", outputSource = "Compile/output")
req1 <- requireScatter()
cwl2 <- cwlWorkflow(requirements = list(req1),
                    inputs = InputParamList(i2),
                    outputs = OutputParamList(o2))
s1 <- cwlStep(id = "Compile", run = Rs,
              In = list(rfile = "cwl_rfiles"),
              scatter = "rfile")
cwl2 <- cwl2 + s1
cwl2
```

Multiple R scripts can be assigned to the workflow inputs and executed.

```{r}
cwl2$cwl_rfiles <- c(r4b$output, r4b$output)
r8 <- runCWL(cwl2, outdir = tempdir())
r8$output
```

## Pipeline plot
The function `plotCWL` can be used to visualize the relationship of
inputs, outputs and the analysis for a tool or pipeline.

```{r}
plotCWL(cwl)
```


