--- 
title: "Bioinformatics tools and pipelines using R and CWL"
author: 
- "Qiang Hu"
- "Qian Liu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "A manual to wrap Bioinformatics tools and workflows with Common Workflow Language using the Rcwl package."
---

# Preface {-}

This short book introduces the R packages, **Rcwl** and **RcwlPipelines**, to improve the way of building, managing and running Bioinformatics tools and pipelines. 

The **Rcwl** package is built on top of the Common Workflow Language (CWL), and provides a simple and user-friendly way to wrap command line tools into data analysis pipelines in R. The **RcwlPipelines** package is a collection of Bioinformatics tools and pipelines based on **Rcwl**.

## R package installation

The **Rcwl** and **RcwlPipelines** packages can be installed from Bioconductor or Github:

```{r eval=FALSE}
BiocManager::install(c("Rcwl", "RcwlPipelines"))
# or the development version
BiocManager::install(c("hubentu/Rcwl", "hubentu/RcwlPipelines"))
```

To load the packages into R session:
```{r}
library(Rcwl)
library(RcwlPipelines)
```

## System requirements

In addition to the R packages, the following tools are required to be installed to run the examples in this book.

* python (>= 2.7)
* cwltool (>= 1.0.2018)
* nodejs

The **cwltool** is the reference implementation of the Common Workflow Language, which is used to run the CWL scripts. The **nodejs** is required when the CWL scripts use JavaScript. You can find instructions to install these tools here:

* <https://github.com/common-workflow-language/cwltool#install>
* <https://nodejs.org>

## Docker

The Docker container simplifies software installation and management, especially for bioinformatics tools/pipelines requiring different runtime environments and library dependencies. A CWL runner can perform this work automatically by pulling the Docker containers and mounting the paths of input files. 

The Docker requirement is optional, as CWL scripts can also be run locally with all the dependencies pre-installed.

## Structure of the book

* Introduction
* Components
* Build CWL workflows
* Run approaches
* Case study

## R session information

The R session information for compiling this mannual is shown below:

```{r}
sessionInfo()
```
